---
title: "Heifer Genotyping Simulations"
author: "Troy Rowan"
date: "8/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(AlphaSimR)
library(tidyverse)

r_to_bif <- function(correlation){
  bif = 1-sqrt((1-correlation^2))
  return(bif)
}

bif_to_r <- function(bif){
  r = sqrt(1-(1-bif)^2)
  return(r)
}
```
## Notes from 08/23/2021 eBeef Meeting
* Using the approach in Newton/Berry likely won't cut it
* Their assumption that the alternative to commercial genomic testing is using a parental average index doesnt' apply to US beef populations
* In reality our commercial selection is all over the place on the female side
    + Bull selection may, however be using GE-EPDs
* Dam selection can be simulated in four ways:
    + Random selection
    + Phenotypic selection
    + Index from a genetic prediction (no genomics)
    + Genomic prediction
* Can we grab accuracies from Angus recalibration update?
* Simulate accuracy by adding random noise to True values to achieve desired reliability
    + Can we just simulate phenotypic selection on traits with different heritabilities?

Questions:
* Do we need to simulate a full dataset first, then use each method to  

#Simulating a founder population for a trait with h2 = 0.25
```{r}
founderPop = runMacs(nInd=100,
                      nChr=10,
                      segSites=500,
                      species = "CATTLE")
#saveRDS(founderPop, "../data/founderpop.RDS")

# BigFounderPop = runMacs(nInd=2000,
#                      nChr=30,
#                      segSites=10000,
#                      species = "CATTLE")
# saveRDS(founderPop, "../data/founderpop.RDS")

#founderPop = readRDS("../data/founderpop.RDS")

SP = SimParam$new(founderPop)
SP$addTraitA(nQtlPerChr=10)
SP$setSexes("yes_sys")
SP$addSnpChip(nSnpPerChr = 100)
SP$addTraitA(nQtlPerChr=rep(10,10))
SP$setVarE(h2 = 0.3)
pheno_pop = ebv_pop = newPop(founderPop)
```

## Simulating 10 generations of selection:
```{r}
gv_pop = rand_pop = ebv_pop = newPop(founderPop)
#This alllows us to track performance over subsequent generations
generations = data.frame(animal = NULL, population = NULL, generation = NULL, BV = NULL, sex = NULL)
#gv_pop = selectCross(pop=gv_pop, nFemale=500, nMale=20, use="gv", nCrosses = 500)
n_crosses = 1000

#This is the selection process that occurs each generation
for(generation in 1:10){
    ans = RRBLUP(ebv_pop, simParam=SP)
    ebv_pop = setEBV(ebv_pop, ans, simParam=SP)
    gv_pop = selectCross(pop=gv_pop, nFemale=150, nMale=20, use="gv", nCrosses = n_crosses) #performs breeding for genetic value selection
    #pheno_pop = selectCross(pop=pheno_pop, nFemale=150, nMale=20, use="pheno", nCrosses = n_crosses) #performs breeding for phenotypic selection
    rand_pop = selectCross(pop=rand_pop, nFemale=150, nMale=20, use="rand", nCrosses = n_crosses)
    ebv_pop = selectCross(pop=ebv_pop, nFemale=150, nMale=20, use="ebv", nCrosses = n_crosses)#performs breeding for random selection
    update = data.frame(animal = seq(1, n_crosses), 
                                     population = c(rep("gv_pop", n_crosses),
                                                    #rep("pheno_pop", n_crosses),
                                                    rep("rand_pop", n_crosses),
                                                    rep("ebv_pop", n_crosses)), 
                                     generation = c(rep(generation, n_crosses),
                                                    #rep(generation, n_crosses),
                                                    rep(generation, n_crosses),
                                                    rep(generation, n_crosses)),
                                     BV = c(gv_pop@gv, rand_pop@gv, ebv_pop@gv),
                        Phenotype = c(gv_pop@pheno, rand_pop@pheno, ebv_pop@gv)
                                       
                        ) 
    generations = bind_rows(generations, update)
    }

generations %>%
    ggplot(aes(generation, BV, color = population))+
    geom_smooth()


for(generation in 1:10){
    gv_pop = selectCross(pop=gv_pop, nFemale=150, nMale=20, use="gv", nCrosses = n_crosses) #performs breeding for genetic value selection
    update = data.frame(animal = seq(1, n_crosses), 
                                     population = rep("gv_pop", n_crosses),
                                     generation = rep(generation, n_crosses),
                                     BV = gv_pop@gv,
                        Phenotype = gv_pop@pheno
                        ) 
    generations = bind_rows(generations, update)
    }

```

## Selecting bulls with EBVs, females via phenotypic selection
* Can we assume that these are coming from two vastly different populations each generation? 
* Does it make sense to have two populations of concurrent selection going on
* Bulls coming from seedstock setting with more strict selection pressure on them. 

```{r}
pop = newPop(founderPop)

ans = RRBLUP(pop, simParam=SP)
pop = setEBV(pop, ans, simParam=SP)

females = selectInd(pop = pop, nInd = 150, sex = "F", use = "pheno")
males = selectInd(pop, nInd = 20, sex = "M", use = "ebv")

makeCross()

gv_pop@gv
```

```{r}
founderPop = quickHaplo(nInd=100, nChr=10, segSites=1000)
# Setting Simulation Parameters
SP = SimParam$new(founderPop)
SP$setSexes("yes_sys")
SP$addTraitA(nQtlPerChr=100)
SP$setVarE(h2=.5)
SP$addSnpChip(10)
# Modeling the Breeding Program
gv_pop = pheno_pop = random_pop = ebv_pop = newPop(founderPop)
ans = RRBLUP(ebv_pop, simParam = SP)
ebv_pop = setEBV(ebv_pop, ans, simParam = SP)
generations = data.frame(animal = NULL, population = NULL, generation = NULL, BV = NULL, sex = NULL)
for(generation in 1:10){
  gv_pop = selectCross(pop=gv_pop, nFemale=45, nMale=5, use="gv", nCrosses = 100)
  pheno_pop = selectCross(pop=pheno_pop, nFemale=45, nMale=5, use="pheno", nCrosses = 100)
  random_pop = selectCross(pop=random_pop, nFemale = 45, nMale = 5, use= "rand", nCrosses = 100)
  ebv_pop = selectCross(pop=ebv_pop, nFemale = 45, nMale = 5, use = "ebv", nCrosses = 100)
  update = data.frame(animal = seq(1, 100),
                      generation = rep(generation, nCrosses = 100),
                      GV = gv_pop@gv,
                      Phenotype = pheno_pop@gv,
                      Random = random_pop@gv,
                      EBV = ebv_pop@gv)
  generations = bind_rows(generations, update)
}
#record the gv, phenotype, and random in a dataframe, set up an empty dataframe.
#Use @ symbol to pull out certain things
#Graphing all together
ggplot(data = generations, aes(x = generation, y = animal)) + ylab(“Genetic Value”) +
  geom_smooth(aes(y = GV), color = “blue”) +
  geom_smooth(aes(y = Phenotype), color = “green”) +
  geom_smooth(aes(y = Random), color = “red”) +
  geom_smooth(aes(y = EBV), color = “orange”)
#1st graph is geom_line, 2nd is geom_point, 3rd is geom_smooth
#geom_smooth is the best line graph, geom_line is super weird and awkward looking
```


###Understanding how to simulate "accuracy"
* Here we're quick-simulating 100,000 animals 
```{r}
founderPop = quickHaplo(nInd=100000, nChr=10, segSites=1000)
SP = SimParam$new(founderPop)
SP$addTraitA(nQtlPerChr=200)
SP$setVarE(h2 = 0.25)
population = newPop(founderPop)

cor(population@gv, population@gv+rnorm(100000,0,1.5), method = "spearman")


```

##Exploring how phenotype and heritability are related
```{r}
pearson_corr = c()
spearman_corr = c()
for (xx in seq(0,1, by =0.01)){
  founderPop = quickHaplo(nInd=100000, nChr=10, segSites=1000)
  SP = SimParam$new(founderPop)
  SP$addTraitA(nQtlPerChr=200)
  SP$setVarE(h2 = xx)
  population = newPop(founderPop)
  pearson_corr = append(pearson_corr, cor(population@gv, population@pheno))
  spearman_corr = append(spearman_corr, cor(population@gv, population@pheno, method = "spearman"))
  }
```

Seeing how adding random noise to genetic values influences correlations between true and "estimated" breeding value
```{r}
founderPop = quickHaplo(nInd=100000, nChr=10, segSites=1000)
SP = SimParam$new(founderPop)
SP$addTraitA(nQtlPerChr=200)
population = newPop(founderPop)
correlations = data.frame(standard_dev=NULL, correlation=NULL)

for (xx in seq(0,10, by = 0.1)){
  update = data.frame(standard_dev = xx, correlation = cor(population@gv, population@gv+rnorm(100000,0,xx), method = "spearman"))
  correlations = bind_rows(correlations, update)
}
correlations %>% 
  rename(correlation = Trait1) %>% 
  dplyr::mutate(reliability = correlation^2,
         BIF = r_to_bif(correlation)) %>% 
  pivot_longer(!standard_dev, names_to = "Measure", values_to = "Accuracy") %>% 
  ggplot(aes(Accuracy, standard_dev, color = Measure))+
  geom_smooth()
```



```{r}
data.frame(h2 = seq(0,1, by =0.01), spearman_corr, pearson_corr) %>% 
  pivot_longer(!h2, names_to = "corr_type", values_to = "correlation") %>% 
  ggplot(aes(h2, correlation, color = corr_type))+
  geom_point()
```


#Heifer selection

###Creating founder population and setting simulation parameters
```{r}
founderPop = quickHaplo(nInd=10000, nChr=10, segSites=1000)
SP = SimParam$new(founderPop)
SP$addTraitA(nQtlPerChr=200)
SP$setSexes("yes_sys")
SP$addSnpChip(nSnpPerChr = 100)
SP$setVarE(h2 = 0.3)


```
### Burn-in simulations and early genetic progress
```{r}
commercial = seedstock = newPop(founderPop)
generations = data.frame(animal = NULL, population = NULL, generation = NULL, BV = NULL, sex = NULL)
rand_ncrosses = 1000
ncrosses = 500

for(generation in -40:-21){
    seedstock = randCross(pop=seedstock, nCrosses = rand_ncrosses)
    commercial = randCross(pop=commercial, nCrosses = rand_ncrosses) #performs breeding for genetic value selection
    update = data.frame(animal = seq(1, seedstock@nInd+commercial@nInd), 
                                     population = c(rep("seedstock", seedstock@nInd),
                                                    rep("commercial", commercial@nInd)), 
                                     generation = rep(generation, seedstock@nInd + commercial@nInd),
                                     BV = c(seedstock@gv, commercial@gv),
                        Phenotype = c(seedstock@pheno, commercial@pheno)
                        ) 
    generations = bind_rows(generations, update)
}

for(generation in -21:-1){
    seedstock@gv <- seedstock@gv+rnorm(seedstock@nInd,0,0.5)
    commercial@gv <- commercial@gv+rnorm(commercial@nInd,0,1.5)
    seedstock = selectCross(pop=seedstock, nFemale = 150, nMale = 150, use = "gv", nCrosses = ncrosses)
    commercial = selectCross(pop=commercial, nFemale = 150, nMale = 150, use = "gv", nCrosses = ncrosses) #performs breeding for genetic value selection
    update = data.frame(animal = seq(1, seedstock@nInd+commercial@nInd), 
                                     population = c(rep("seedstock", seedstock@nInd),
                                                    rep("commercial", commercial@nInd)), 
                                     generation = rep(generation, seedstock@nInd + commercial@nInd),
                                     BV = c(seedstock@gv, commercial@gv),
                        Phenotype = c(seedstock@pheno, commercial@pheno)
                        ) 
    generations = bind_rows(generations, update)
}

```

Genetic progress of commercial vs seedstock over burn-in phase
```{r}
generations %>% 
  ggplot(aes(x = generation, y = BV, color = population))+
  geom_smooth()
```
## Simulating commercial genomics
Non-overlapping generations (Each female is crossed 5 times)
```{r}
commercial = seedstock = newPop(founderPop)

generations = data.frame(animal = NULL, population = NULL, generation = NULL, BV = NULL, sex = NULL)

rand_ncrosses = 1000
ncrosses = 500

commercial_pheno = commercial
commercial_genomic = commercial
for (generation in 0:20){
  seedstock@gv <- seedstock@gv+rnorm(seedstock@nInd,0,0.5)
  commercial_pheno@gv <- commercial@gv+rnorm(commercial@nInd,0,1.5)
  commercial_genomic@gv <- commercial@gv+rnorm(commercial@nInd,0,0.5)
  
  seedstock = selectCross(pop = seedstock, nFemale = 1000, nMale = 50, use = "gv", nCrosses = 5000)
  
  bulls = selectInd(seedstock, nInd = 250, use = "gv", sex = "M")
  heifers_genomic = selectInd(commercial, nInd = 1000, use = "gv", sex = "F")
  heifers_pheno = selectInd(commercial, nInd = 1000, use = "gv", sex = "F")
  
  commercial_pheno = randCross2(females = heifers_pheno, males = bulls, nProgeny = 5, balance = TRUE, nCrosses = 5000)
  commercial_genomic = randCross2(females = heifers_genomic, males = bulls, nProgeny = 5, balance = TRUE, nCrosses = 5000)
  
  update = data.frame(animal = seq(1, seedstock@nInd+commercial_pheno@nInd, commercial_genomic@nInd), 
                                     population = c(rep("seedstock", seedstock@nInd),
                                                    rep("commercial_phenotypic", commercial_pheno@nInd), rep("commercial_genomic", commercial_genomic@nInd)), 
                                     generation = rep(generation, seedstock@nInd + commercial_pheno@nInd + commercial_genomic@nInd),
                                     BV = c(seedstock@gv, commercial_pheno@gv, commercial_genomic@gv),
                        Phenotype = c(seedstock@pheno, commercial_pheno@pheno, commercial_genomic@pheno)
                        )
  
  generations = bind_rows(generations, update)
}
```

