---
title: "Heifer Genotyping Simulations"
author: "Troy Rowan"
date: "8/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(AlphaSimR)
```
## Notes from 08/23/2021 eBeef Meeting
* Using the appraoch in Newton/Berry likely won't cut it
* Their assumption that the alternative to commercial genomic testing is using a parental average index doesnt' apply to US beef populations
* In reality our commercial selection is all over the palce on the female side
    + Bull selection may, however be using GE-EPDs
* Dam selection can be simulated in four ways:
    + Random selection
    + Phenotypic selection
    + Index from a genetic prediction (no genomics)
    + Genomic prediction
* Can we grab accuracies from Angus recalibration update?
* Simulate accuracy by adding random noise to True values to achieve desired reliability
    + Can we just simulate phenotypic selection on traits with different heritabilities?

Questions:
* Do we need to simulate a full dataset first, then use each method to  

#Simulating a founder population for a trait with h2 = 0.25
```{r}
# founderPop = runMacs(nInd=1000,
#                      nChr=10,
#                      segSites=5000,
#                      species = "CATTLE")
#saveRDS(founderPop, "../data/founderpop.RDS")

BigFounderPop = runMacs(nInd=2000,
                     nChr=30,
                     segSites=10000,
                     species = "CATTLE")
saveRDS(founderPop, "../data/founderpop.RDS")

founderPop = readRDS("../data/founderpop.RDS")

SP = SimParam$new(founderPop)
SP$addTraitA(nQtlPerChr=1000)
SP$setSexes("yes_sys")
SP$addSnpChip(nSnpPerChr = 500)
SP$setVarE(h2 = 0.25)

```

## Simulating 10 generations of selection:
```{r}
pheno_pop = gv_pop = rand_pop = ebv_pop = newPop(founderPop)

generations = data.frame(animal = NULL, population = NULL, generation = NULL, BV = NULL, sex = NULL)
#gv_pop = selectCross(pop=gv_pop, nFemale=500, nMale=20, use="gv", nCrosses = 500)
n_crosses = 1000

for(generation in 1:10){
    ans = RRBLUP(ebv_pop, simParam=SP)
    ebv_pop = setEBV(ebv_pop, ans, simParam=SP)
    gv_pop = selectCross(pop=gv_pop, nFemale=150, nMale=20, use="gv", nCrosses = n_crosses) #performs breeding for genetic value selection
    pheno_pop = selectCross(pop=pheno_pop, nFemale=150, nMale=20, use="pheno", nCrosses = n_crosses) #performs breeding for phenotypic selection
    rand_pop = selectCross(pop=rand_pop, nFemale=150, nMale=20, use="rand", nCrosses = n_crosses)
    ebv_pop = selectCross(pop=ebv_pop, nFemale=150, nMale=20, use="ebv", nCrosses = n_crosses)#performs breeding for random selection
    update = data.frame(animal = seq(1, n_crosses), 
                                     population = c(rep("gv_pop", n_crosses),
                                                    rep("pheno_pop", n_crosses),
                                                    rep("rand_pop", n_crosses),
                                                    rep("ebv_pop", n_crosses)), 
                                     generation = c(rep(generation, n_crosses),
                                                    rep(generation, n_crosses),
                                                    rep(generation, n_crosses),
                                                    rep(generation, n_crosses)),
                                     BV = c(gv_pop@gv, pheno_pop@gv, rand_pop@gv, ebv_pop@gv),
                        Phenotype = c(gv_pop@pheno, pheno_pop@pheno, rand_pop@pheno, ebv_pop@gv)
                                       
                        ) 
    generations = bind_rows(generations, update)
    }

generations %>%
    ggplot(aes(generation, BV, color = population))+
    geom_smooth()

```

## Selecting bulls with EBVs, females via phenotypic selection
```{r}
pop = newPop(founderPop)

ans = RRBLUP(pop, simParam=SP)
pop = setEBV(pop, ans, simParam=SP)

females = selectInd(pop = pop, nInd = 150, sex = "F", use = "pheno")
males = selectInd(pop, nInd = 150)

makeCross()

```

